---
title: "Manuscript 1 code"
author: "Noam Altman-Kurosaki, Callie Stephenson"
date: "2024-09-11"
output: html_document
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(here)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(vegan)
library(glmmTMB)
library(car)
library(effects)
library(DHARMa)
library(stringr)
library(MuMIn)
library(codyn)
library(nlme)
library(mgcv)
library(emmeans)
```

## functions

```{r functions}
# coefficient of variation
CV<-function(x){
  return(sd(x,na.rm=T)/mean(x,na.rm=T))
}

inv_logit <- function(x){
  exp(x)/(1+exp(x))
}

# smithson verkuilen transformation for 0-1 continuous data
sv_trans <- function(prop, s = 0.000005){
  (prop*(length(prop) - 1) + s)/length(prop)
  # where prop is the vector of the proportional value you're transforming, 
  # N is the sample size, which is specified by taking the number of rows/observations from a given dataframe,
  # and s is a small offset 
}

se <- function(x){
  sd(x)/sqrt(n())
}

```


## Figure formatting
```{r figure formatting info}
habitat_colours <- c("Fringing" = "#D81B60",
                     "Backreef" = "#FFC107",
                     "Forereef 10m" = "#1E88E5",
                     "Forereef 17m" = "#004D40")

functional_group_colours <- c("Coral" = "darkmagenta",
                         "CCA" = "deeppink3",
                         "Macroalgae" = "chartreuse4",
                         "Turf" =  "#004D40")

```

# read and clean data

```{r read and clean data}
data <- read.csv(here::here("data", "allbenthicdata_cleaned_colnames.csv"), stringsAsFactors = F) 


not_algae <- c("ascidian", "bare_space", "coral", "coral_rubble", "corallimorpharia", "heteractis_sp", "millepora_platyphylla", "no_data", "sand", "sarcophyton_sp", "shell_debris", "soft_coral", "sponge", "tridacna_sp", "zooxanthellae", "row_sums")



algae <- data[ , -which(names(data) %in% not_algae)]



macroalgae <- algae %>% 
  select(-algal_turf, -crustose_corallines, -cyanophyta, -damselfish_turf, -lithophyllum_kotschyanum,
         -neogoniolithon_brassica_florida, -sporolithon_sp)

meta_cols <- c("year", "location", "site", "habitat", "site_habitat", "transect", "quadrat")

meta_df <- algae[,which(names(algae) %in% meta_cols)]

algae_species_cols <- algae[,-which(names(algae) %in% meta_cols)]

macro_species_cols <- macroalgae[,-which(names(macroalgae) %in% meta_cols)]


```


```{r site macroalgal data}
site_macro <- macroalgae %>%
  dplyr::group_by(year, site, habitat,site_habitat) %>%
  dplyr::summarize(across(acanthophora_spicifera:valonia_ventricosa, \(x) mean(x, na.rm = TRUE)))

site_meta_cols <- c("year", "site", "habitat", "site_habitat")
site_meta_df <- site_macro[,which(names(site_macro) %in% meta_cols)]
```

# Figure 1

## setup

```{r cover plot setup}
cover_df <- data.frame(
  year = data$year,
  habitat = data$habitat,
  site = data$site,
  location = data$location,
  coral = data$coral,
  cca = data$crustose_corallines + data$lithophyllum_kotschyanum + data$neogoniolithon_brassica_florida + data$sporolithon_sp,
  macroalgae = rowSums(macro_species_cols),
  turf = data$algal_turf  + data$damselfish_turf, 
  sand = data$sand,
  coral_rubble = data$coral_rubble,
  cyanophyta = data$cyanophyta,
  other = data$ascidian + data$corallimorpharia + data$heteractis_sp + data$millepora_platyphylla + data$sarcophyton_sp + data$shell_debris +
    data$soft_coral + data$sponge + data$tridacna_sp + data$zooxanthellae + data$no_data + data$bare_space
)

cover_df$total <- rowSums(cover_df[,-which(names(cover_df) %in% meta_cols)])

cover_means <- cover_df %>%
  group_by(year, habitat) %>%
  summarise(macroalgae = mean(macroalgae),
            coral = mean(coral),
            turf = mean(turf),
            cca = mean(cca),
            sand = mean(sand),
            other = mean(other)
            )

cover_se <- cover_df %>%
  group_by(year, habitat) %>%
  summarise(macroalgae = se(macroalgae),
            coral = se(coral),
            turf = se(turf),
            cca = se(cca),
            sand = se(sand),
            other = se(other)
            )
```

```{r cover long}
cover_long_means <- pivot_longer(
  data = cover_means,
  cols = 3:8,
  names_to = "functional_group",
  values_to = "percent_cover"
)

cover_long_se <- pivot_longer(
  data = cover_se,
  cols = 3:8,
  names_to = "functional_group",
  values_to = "se"
)

cover_long <- left_join(cover_long_means,cover_long_se)

#make levels for habitat for plotting
cover_long$habitat <- factor(cover_long$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))

cover_long$functional_group <- factor(cover_long$functional_group, 
                             levels = c("macroalgae", "coral", "turf","cca", "sand", "other"),
                             labels = c("Macroalgae", "Coral", "Turf", "CCA", "Sand", "Other"))
```

```{r simplified cover - remove}

# simplified_cover <- data.frame(
#   year = cover_df$year,
#   habitat = cover_df$habitat,
#   site = cover_df$site,
#   location = cover_df$location,
#   low_lying_algae = cover_df$cca + cover_df$turf + cover_df$cyanophyta,
#   coral = cover_df$coral,
#   macroalgae = cover_df$macroalgae
# )
# 
# simplified_cover_means <- simplified_cover %>%
#   group_by(year, habitat) %>%
#   summarise(macroalgae_mean = mean(macroalgae),
#             low_lying_algae_mean = mean(low_lying_algae),
#             coral_mean = mean(coral),
#             macroalgae_se = sd(macroalgae)/sqrt(n()),
#             low_lying_algae_se = sd(low_lying_algae)/sqrt(n()),
#             coral_se = sd(coral)/sqrt(n())
#             )
# 
# simplified_cover_long <- pivot_longer(
#   data = simplified_cover_means,
#   cols = 3:5,
#   names_to = "functional_group",
#   values_to = "percent_cover"
# )
# 
# simplified_cover_long_se <- pivot_longer(
#   data = simplified_cover_means,
#   cols = 6:8,
#   names_to = "functional_group",
#   values_to = "se"
# )
# simplified_cover_long$se <- simplified_cover_long_se$se

```

```{r}
figure_1_data <- subset(cover_long, functional_group != "Sand" & functional_group != "Other")
```

## plot


```{r figure 1, fig.height = 10, fig.width=14, warning=FALSE}
(figure_1 <- ggplot() +
  geom_line(data = figure_1_data, aes(x = year, y = percent_cover, colour = functional_group), linewidth = 1) +
  geom_ribbon(data = figure_1_data, aes(x = year, ymin = percent_cover - se, ymax = percent_cover + se, fill = functional_group), alpha = 0.3) +
  facet_wrap(~ habitat) +
  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
  annotate("segment", x = 2006, xend = 2010, y = 75, yend = 75, linewidth = 2, colour = "grey") +
  scale_colour_manual(values = functional_group_colours) +
  scale_fill_manual(values = functional_group_colours) +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(cover_long$year)) +
   labs(y = "Percent Cover \n",
        x = "Year") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

#ggsave(filename = "output/figure_1.png", figure_1, height = 10, width = 14)
```

# Figure 2

## Create Data Frames
```{r create data}
alpha_diversity_quad_macro <-
  data.frame(
    year = macroalgae$year,
    location = macroalgae$location,
    shannon = diversity(macroalgae[ , -which(names(macroalgae) %in% meta_cols)], "shannon"),
    richness = rowSums(macroalgae[ , -which(names(macroalgae) %in% meta_cols)] > 0)
  )

alpha_diversity_quad_macro <- merge(meta_df, alpha_diversity_quad_macro, by = c("location", "year"))
```

```{r site level richness df}
alpha_diversity_site_macro <-
  data.frame(
    year = site_macro$year,
    site_habitat = site_macro$site_habitat,
    shannon = diversity(site_macro[ , -which(names(site_macro) %in% meta_cols)], "shannon"),
    richness = rowSums(site_macro[ , -which(names(site_macro) %in% meta_cols)] > 0)
  ) %>% 
  left_join(site_meta_df, alpha_diversity_site_macro, by = c("site_habitat", "year"))
```


# Figure 2

## Figure 2a

### prep

```{r richness df}
quad_richness_df <- alpha_diversity_quad_macro %>% 
  group_by(year, habitat) %>% 
  summarise(richness_mean = mean(richness),
            richness_se = se(richness))
```

```{r richness factors}
quad_richness_df$habitat <- factor(quad_richness_df$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
```

### plot
```{r figure 2}
(figure_2a <- ggplot() +
  geom_line(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
  geom_ribbon(data = quad_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se, fill = habitat), alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
  annotate("segment", x = 2006, xend = 2010, y = 2, yend = 2, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(quad_richness_df$year)) +
   labs(y = "Macroalgal Richness \n",
        x = "Year",
        title = "a. Quadrat-level Richness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size=18), 
        plot.title = element_text(size=18),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

ggsave(filename = "output/figure_2a.png", figure_2a, height = 10, width = 14)
```

## Figure 2b
### prep

```{r richness df}
site_richness_df <- alpha_diversity_site_macro %>% 
  group_by(year, habitat) %>% 
  summarise(richness_mean = mean(richness),
            richness_se = se(richness))

```

```{r richness factors}
site_richness_df$habitat <- factor(site_richness_df$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))


```

```{r}
(figure_2b <- ggplot() +
  geom_line(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
  geom_ribbon(data = site_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se, fill = habitat), alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
  annotate("segment", x = 2006, xend = 2010, y = 10, yend = 10, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(site_richness_df$year)) +
   labs(y = "Macroalgal Richness \n",
        x = "Year",
        title = "b. Site-level Richness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size=18), 
        plot.title = element_text(size=18),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

ggsave(filename = "output/figure_2b.png", figure_2b, height = 10, width = 14)
```


# Figure 3

## set up data for model
```{r quad data setup}
macroalgal_cover <- cover_df[,c("year", "habitat", "site", "location", "macroalgae")]

alpha_diversity_quad_macro_with_cover <- left_join(alpha_diversity_quad_macro, macroalgal_cover,by = join_by(location, year, site, habitat)) %>% 
  mutate(macroalgal_prop_cover = macroalgae / 100) %>% 
  select(-macroalgae)
```

```{r site data setup}
alpha_diversity_site_macro_with_cover <- alpha_diversity_site_macro %>% 
  mutate(percent_cover = rowSums(site_macro[,-which(names(site_macro)  %in% colnames(site_meta_df))])) %>% 
  mutate(prop_cover = percent_cover / 100) %>% 
  select(-percent_cover) %>% 
  mutate(cover_trans = sv_trans(prop_cover))
```


## models

### quad

```{r figure 3a model}
alpha_diversity_quad_macro_with_cover$cover_trans <- sv_trans(alpha_diversity_quad_macro_with_cover$macroalgal_prop_cover)

cover_mod <- glmmTMB(cover_trans ~ richness*habitat + (1|site/location) + (1|year), family = beta_family(), data = alpha_diversity_quad_macro_with_cover)
summary(cover_mod)
hist(residuals(cover_mod)) # looks fine
plot(residuals(cover_mod) ~ fitted(cover_mod)) # negative trend but it can't be less than 0 so not too concerned. some wave pattern
r.squaredGLMM(cover_mod) # 0.631, 0.650
emtrends(cover_mod, pairwise ~ habitat, var = "richness") # backreef not different from forereef, forereef also not different
```

### site
```{r site mod}
site_cover_mod <- glmmTMB(cover_trans ~ richness*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = alpha_diversity_site_macro_with_cover)
summary(site_cover_mod)
hist(residuals(site_cover_mod)) # some positive skew, not the worst thing ever
plot(residuals(site_cover_mod) ~ fitted(site_cover_mod)) # a little heteroscedastic but can't be less than 0
r.squaredGLMM(site_cover_mod) # 0.2212534, 0.5990825
emtrends(site_cover_mod, pairwise ~ habitat, var = "richness") # no difference between the habitats
```



## figure 3a plot setup

```{r figure 3a setup}
diversity_ranges_quad <- alpha_diversity_quad_macro_with_cover %>%
  group_by(habitat) %>%
  summarise(min_richness = min(richness),
            max_richness = max(richness),
            min_cover= min(cover_trans),
            max_cover = max(cover_trans))

cover_emm <- emmip(cover_mod,
                        habitat ~ richness,
                        at = list(richness = seq(0,10,1)), plotit = F, CIs = T)

filtered_cover_effects <- cover_emm %>%
  left_join(diversity_ranges_quad, by = c("habitat")) %>%
  filter(richness >= min_richness & richness <= max_richness)

alpha_diversity_quad_macro_with_cover$habitat <- factor(alpha_diversity_quad_macro_with_cover$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))


filtered_cover_effects$habitat <- factor(filtered_cover_effects$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
```

## figure 3a plot

```{r figure 3a plot}
(figure_3a <- 
  ggplot() +
  geom_point(data = alpha_diversity_quad_macro_with_cover,
             aes(x = richness, y = cover_trans, colour = habitat), size = 1, alpha = 0.005) + 
  geom_line(data = filtered_cover_effects, aes(x = richness, y = inv_logit(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_cover_effects, aes(x = richness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Cover (Proprtion)\n", x = "Macroalgal Richness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )
ggsave(filename = "output/figure_3a.png", figure_3a, height = 10, width = 14)
```

## Figure 3d

### setup
```{r}
diversity_ranges_site <- alpha_diversity_site_macro_with_cover %>%
  group_by(habitat) %>%
  summarise(min_richness = min(richness),
            max_richness = max(richness),
            min_cover= min(cover_trans),
            max_cover = max(cover_trans))

site_cover_emm <- emmip(site_cover_mod,
                        habitat ~ richness,
                        at = list(richness = seq(0,17,1)), plotit = F, CIs = T)

filtered_site_cover_effects <- site_cover_emm %>%
  left_join(diversity_ranges_site, by = c("habitat")) %>%
  filter(richness >= min_richness & richness <= max_richness)

alpha_diversity_site_macro_with_cover$habitat <- factor(alpha_diversity_site_macro_with_cover$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))


filtered_site_cover_effects$habitat <- factor(filtered_site_cover_effects$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
```

### plot

```{r}
(figure_3d <- 
  ggplot() +
  geom_point(data = alpha_diversity_site_macro_with_cover,
             aes(x = richness, y = cover_trans, colour = habitat), size = 1, alpha = 0.3) + 
  geom_line(data = filtered_site_cover_effects, aes(x = richness, y = inv_logit(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_site_cover_effects, aes(x = richness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Cover (Proprtion)\n", x = "Macroalgal Richness", title = "3d.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        plot.title = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )
ggsave(filename = "output/figure_3d.png", figure_3d, height = 10, width = 14)
```


# Figure 4

## Synchrony setup
```{r calculate syncrony}
# make er long
macro_long_data <- pivot_longer(
  data = macroalgae,
  cols = grep("acanthophora_spicifera", colnames(macroalgae)):grep("valonia_ventricosa", colnames(macroalgae)),
  names_to = "taxa",
  values_to = "percent_cover") %>% 
  mutate(prop_cover = percent_cover/100) %>% 
  select(-percent_cover)

# write_csv(macro_long_data, here("Data", "macro_long_data.csv"))

# CALCULATE SYNCHRONY CODE
lm_synchrony <- codyn::synchrony(df = macro_long_data,
                                   time.var = "year",
                                   species.var = "taxa",
                                   abundance.var = "prop_cover",
                                   metric = "Loreau",
                                   replicate.var = "location")


diversity_stability <- alpha_diversity_quad_macro_with_cover %>%
  dplyr::group_by(location, habitat, site) %>%
  dplyr::summarise(
            richness_mean = mean(richness), 
            shannon_mean = mean(shannon), 
            cover_mean = mean(macroalgal_prop_cover),
            cover_cv = CV(macroalgal_prop_cover),
            cover_stability = 1/CV(macroalgal_prop_cover)) 

diversity_stability_synchrony <- left_join(diversity_stability, lm_synchrony, by = join_by(location))
diversity_stability_synchrony$synchrony_trans <- sv_trans(diversity_stability_synchrony$synchrony)
diversity_stability_synchrony <- na.omit(diversity_stability_synchrony) # remove quads that never had algae --> NaN calcs


dss_ranges<- diversity_stability_synchrony %>%
  group_by(habitat) %>%
  summarise(min_richness = min(richness_mean),
            max_richness = max(richness_mean),
            min_stability= min(cover_stability),
            max_stability = max(cover_stability),
            min_synchrony = min(synchrony_trans),
            max_synchrony = max(synchrony_trans))
```

## figure 4a model - richness stability

```{r richness synchrony mod}
rich_stab_mod <- glmmTMB(cover_stability~ richness_mean*habitat + (1|site), family = Gamma(link = "log"), data = diversity_stability_synchrony)
summary(rich_stab_mod)
hist(residuals(rich_stab_mod)) # looks good
plot(residuals(rich_stab_mod) ~ fitted(rich_stab_mod)) # heteroskedastic - bring to group
r.squaredGLMM(rich_stab_mod) # 0.803, 0.808
emtrends(rich_stab_mod, pairwise ~ habitat, var = "richness_mean") # fringing not diff from outer17, backreef not different from outer10
```

```{r figure 4a setup}


rich_stab_emm <- emmip(rich_stab_mod,
                        habitat ~ richness_mean,
                        at = list(richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)

filtered_rich_stab_effects <- rich_stab_emm %>%
  left_join(dss_ranges, by = c("habitat")) %>%
  filter(richness_mean >= min_richness & richness_mean <= max_richness)


```


## plot figure 4a

```{r figure 4a}
(figure_4a <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony,
             aes(x = richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_rich_stab_effects, aes(x = richness_mean, y = exp(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_rich_stab_effects, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Stability\n", x = "Macroalgal Richness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

ggsave(filename = "output/figure_4a.png", figure_4a, height = 10, width = 14)
```

## figure 4a no habitat

```{r richness stability no hab}
rich_stab_mod_2 <- glmmTMB(cover_stability~ richness_mean + (1|site), family = Gamma(link = "log"), data = diversity_stability_synchrony)
summary(rich_stab_mod_2)
hist(residuals(rich_stab_mod_2)) # a chunkier version
plot(residuals(rich_stab_mod_2) ~ fitted(rich_stab_mod_2)) # heteroskedastic - bring to group
r.squaredGLMM(rich_stab_mod_2) # 0.750, 0.757


```

```{r 4a no hab setup}
rich_stab_emm_2 <-as.data.frame(emmeans(rich_stab_mod_2, ~richness_mean,
                                     at = list(richness_mean = seq(min(diversity_stability_synchrony$richness_mean),
                                                            max(diversity_stability_synchrony$richness_mean),
                                                            0.01)))) 

```

## figure 4a no habitat plot

```{r figure 4a no habitat}
(figure_4a_no_hab <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony,
             aes(x = richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_rich_stab_effects, aes(x = richness_mean, y = exp(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_rich_stab_effects, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  geom_line(data = rich_stab_emm_2, aes(x = richness_mean, y = exp(emmean)), size = 2, colour = "black") +
  geom_ribbon(data = rich_stab_emm_2, aes(x = richness_mean, ymin = exp(asymp.LCL), ymax = exp(asymp.UCL)), fill = "black",
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Stability\n", x = "Macroalgal Richness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

ggsave(filename = "output/figure_4a_with_no_hab.png", figure_4a_no_hab, height = 10, width = 14)
```

## figure 4c model -  synchrony staability

```{r synchrony staability mod}
synch_stab_mod <- glmmTMB(cover_stability ~ synchrony*habitat + (1|site), family = Gamma("log"), data = diversity_stability_synchrony)
summary(synch_stab_mod)
hist(residuals(synch_stab_mod)) # looks good
plot(residuals(synch_stab_mod) ~ fitted(synch_stab_mod)) # same issue as always
r.squaredGLMM(synch_stab_mod) # 0.63, 0.71
emtrends(synch_stab_mod, pairwise ~ habitat, var = "synchrony") # fringing and backreef not different
```

```{r figure 4c setup}


synch_stab_emm <- emmip(synch_stab_mod,
                        habitat ~ synchrony,
                        at = list(synchrony = seq(0,1,0.01)), plotit = F, CIs = T)

filtered_synch_stab_effects <- synch_stab_emm %>%
  left_join(dss_ranges, by = c("habitat")) %>%
  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)


```

## plot figure 4c

```{r figure 4c}
(figure_4c <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony,
             aes(x = synchrony, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_synch_stab_effects, aes(x = synchrony, y = exp(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_synch_stab_effects, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(x = "Synchrony", y = "Stability\n") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )
ggsave(filename = "output/figure_4c.png", figure_4c, height = 10, width = 14)
```



# Figure 3c

## synchrony and cover model

```{r synchrony cover mod}
synch_cov_mod <- glmmTMB(cover_mean ~ synchrony*habitat + (1|site), family = beta_family(), data = diversity_stability_synchrony)
summary(synch_cov_mod)
hist(residuals(synch_cov_mod)) # slight skew
plot(residuals(synch_cov_mod) ~ fitted(synch_cov_mod)) # same issues as always
emtrends(synch_cov_mod, pairwise ~ habitat, var = "synchrony") # forereef 17 distinct from all other habitats
```

```{r synch cov plot setup}
synch_cov_emm <- emmip(synch_cov_mod,
                        habitat ~ synchrony,
                        at = list(synchrony = seq(0,1,0.01)), plotit = F, CIs = T)

filtered_synch_cov_effects <- synch_cov_emm %>%
  left_join(dss_ranges, by = c("habitat")) %>%
  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
```

## plot

```{r synchrony cover plot}
(figure_3c <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony,
             aes(x = synchrony, y = cover_mean, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_synch_cov_effects, aes(x = synchrony, y = exp(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_synch_cov_effects, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(x = "Synchrony", y = "Cover\n") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )


ggsave(filename = "output/figure_3c.png", figure_3c, height = 10, width = 14)
```




## Synchrony setup site level
```{r calculate syncrony site_level}
# make er long
macro_long_data_site <- pivot_longer(
  data = site_macro,
  cols = grep("acanthophora_spicifera", colnames(site_macro)):grep("valonia_ventricosa", colnames(site_macro)),
  names_to = "taxa",
  values_to = "percent_cover") %>% 
  mutate(prop_cover = percent_cover/100) %>% 
  select(-percent_cover)

# write_csv(macro_long_data, here("Data", "macro_long_data.csv"))

# CALCULATE SYNCHRONY CODE
lm_synchrony_site <- codyn::synchrony(df = macro_long_data_site,
                                   time.var = "year",
                                   species.var = "taxa",
                                   abundance.var = "prop_cover",
                                   metric = "Loreau",
                                   replicate.var = "site_habitat")


diversity_stability_site <- alpha_diversity_site_macro_with_cover %>%
  dplyr::group_by(site_habitat, habitat, site) %>%
  dplyr::summarise(
            richness_mean = mean(richness), 
            shannon_mean = mean(shannon), 
            cover_mean = mean(prop_cover),
            cover_cv = CV(prop_cover),
            cover_stability = 1/CV(prop_cover)) 

diversity_stability_synchrony_site <- left_join(diversity_stability_site, lm_synchrony_site, by = join_by(site_habitat))
diversity_stability_synchrony_site$synchrony_trans <- sv_trans(diversity_stability_synchrony_site$synchrony)
# diversity_stability_synchrony_site <- na.omit(diversity_stability_synchrony_site) # remove quads that never had algae --> NaN calcs


dss_ranges_site<- diversity_stability_synchrony_site %>%
  group_by(habitat) %>%
  summarise(min_richness = min(richness_mean),
            max_richness = max(richness_mean),
            min_stability= min(cover_stability),
            max_stability = max(cover_stability),
            min_synchrony = min(synchrony_trans),
            max_synchrony = max(synchrony_trans))
```


## figure 4d model - richness stability

```{r richness synchrony mod}
rich_stab_mod_site <- glmmTMB(cover_stability~ richness_mean*habitat, family = Gamma(link = "log"), data = diversity_stability_synchrony_site)
summary(rich_stab_mod_site)
hist(residuals(rich_stab_mod_site)) # blocky but fine
plot(residuals(rich_stab_mod_site) ~ fitted(rich_stab_mod_site)) # heteroskedastic - bring to group
r.squaredGLMM(rich_stab_mod_site) # 0.671
emtrends(rich_stab_mod_site, pairwise ~ habitat, var = "richness_mean") # no differences across habitats
```

```{r figure 4d setup}


rich_stab_emm_site <- emmip(rich_stab_mod_site,
                        habitat ~ richness_mean,
                        at = list(richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)

filtered_rich_stab_effects_site <- rich_stab_emm_site %>%
  left_join(dss_ranges_site, by = c("habitat")) %>%
  filter(richness_mean >= min_richness & richness_mean <= max_richness)


```


## plot figure 4d

```{r figure 4d}
(figure_4d <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony_site,
             aes(x = richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_rich_stab_effects_site, aes(x = richness_mean, y = exp(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_rich_stab_effects_site, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Stability\n", x = "Macroalgal Richness", title = "4d.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        plot.title = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

ggsave(filename = "output/figure_4d.png", figure_4d, height = 10, width = 14)
```


## figure 4f

```{r synchrony staability mod site}
synch_stab_mod_site <- glmmTMB(cover_stability ~ synchrony*habitat, family = Gamma("log"), data = diversity_stability_synchrony_site)
summary(synch_stab_mod_site)
hist(residuals(synch_stab_mod_site)) # almost uniform
plot(residuals(synch_stab_mod_site) ~ fitted(synch_stab_mod_site)) # same issue as always
r.squaredGLMM(synch_stab_mod_site) # 0.678
emtrends(synch_stab_mod_site, pairwise ~ habitat, var = "synchrony") # fringing and backreef are different from forereef 17
```

```{r figure 4f setup}


synch_stab_emm_site <- emmip(synch_stab_mod_site,
                        habitat ~ synchrony,
                        at = list(synchrony = seq(0,1,0.01)), plotit = F, CIs = T)

filtered_synch_stab_effects_site <- synch_stab_emm_site %>%
  left_join(dss_ranges_site, by = c("habitat")) %>%
  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)


```

## plot figure 4f

```{r figure 4f}
(figure_4f <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony_site,
             aes(x = synchrony, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_synch_stab_effects_site, aes(x = synchrony, y = exp(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_synch_stab_effects_site, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(x = "Synchrony", y = "Stability\n", main = "4f.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18),
        plot.title = element_text(size=18),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_text(size = rel(1)),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )
ggsave(filename = "output/figure_4f.png", figure_4f, height = 10, width = 14)
```

# figure 3f

## model

```{r synchrony cover mod}
synch_cov_mod_site <- glmmTMB(cover_mean ~ synchrony*habitat, family = beta_family(), data = diversity_stability_synchrony_site)
summary(synch_cov_mod_site)
hist(residuals(synch_cov_mod_site)) # slight skew
plot(residuals(synch_cov_mod_site) ~ fitted(synch_cov_mod_site)) # looks fine
emtrends(synch_cov_mod_site, pairwise ~ habitat, var = "synchrony") # forereef 17 different from backreef
```

```{r synch cov site plot setup}
synch_cov_emm_site<- emmip(synch_cov_mod_site,
                        habitat ~ synchrony,
                        at = list(synchrony = seq(0,1,0.01)), plotit = F, CIs = T)

filtered_synch_cov_effects_site <- synch_cov_emm_site %>%
  left_join(dss_ranges_site, by = c("habitat")) %>%
  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
```

## plot figure 3f

```{r synchrony cover plot site}
(figure_3f <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony_site,
             aes(x = synchrony, y = cover_mean, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_synch_cov_effects_site, aes(x = synchrony, y = exp(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_synch_cov_effects_site, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(x = "Synchrony", y = "Cover\n", title = "3f.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        plot.title = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )


ggsave(filename = "output/figure_3f.png", figure_3f, height = 10, width = 14)
```