---
title: "Manuscript 1 code"
author: "Noam Altman-Kurosaki, Callie Stephenson"
date: "2024-09-11"
output: html_document
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
library(here)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(vegan)
library(glmmTMB)
library(car)
library(effects)
library(DHARMa)
library(stringr)
library(MuMIn)
library(codyn)
library(nlme)
library(mgcv)
library(emmeans)
library(dplyr)
```

## Functions
```{r}
source("R/functions.R")
```

## Figure formatting
```{r figure formatting info}
mytheme <- theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size=18), 
        panel.background = element_rect(fill = "white"), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) 

habitat_colours <- c("Fringing" = "#D81B60",
                     "Backreef" = "#FFC107",
                     "Forereef 10m" = "#1E88E5",
                     "Forereef 17m" = "#004D40")

functional_group_colours <- c("Coral" = "darkmagenta",
                         "CCA" = "deeppink3",
                         "Macroalgae" = "chartreuse4",
                         "Turf" =  "#004D40")

```

# read and clean data

```{r read and clean data}
data <- read.csv(here::here("data", "allbenthicdata_cleaned_colnames.csv"), stringsAsFactors = F) 

### if we want to make a missing or faulty data table
# length(unique(data$location))
# expected_data <- expand.grid(year = seq(2007,2023,1), location = unique(data$location))
# missing_quadrats <- expected_data %>%
#   anti_join(data, by = c("year", "location"))
###

# remove remaining backreef and fringing 2020 data because of covid extrapolation
filtered_data <- data %>%
  filter(!(year == 2020 & habitat %in% c("fringing", "backreef")))

#THIS IS THE BIG CHANGE HERE: 9/19
filtered_data$habitat <- factor(filtered_data$habitat, 
                             levels = c("fringing", "backreef", "outer_10","outer_17"),
                             labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))

#DID NOT TOUCH BELOW:
not_algae <- c("ascidian", "bare_space", "coral", "coral_rubble", "corallimorpharia", "heteractis_sp", "millepora_platyphylla", "no_data", "sand", "sarcophyton_sp", "shell_debris", "soft_coral", "sponge", "tridacna_sp", "zooxanthellae", "row_sums")



algae <- filtered_data[ , -which(names(filtered_data) %in% not_algae)]



macroalgae <- algae %>% 
  select(-algal_turf, -crustose_corallines, -cyanophyta, -damselfish_turf, -lithophyllum_kotschyanum,
         -neogoniolithon_brassica_florida, -sporolithon_sp)

meta_cols <- c("year", "location", "site", "habitat", "site_habitat", "transect", "quadrat")

meta_df <- algae[,which(names(algae) %in% meta_cols)]

algae_species_cols <- algae[,-which(names(algae) %in% meta_cols)]

macro_species_cols <- macroalgae[,-which(names(macroalgae) %in% meta_cols)]
```


```{r site macroalgal data}
site_macro <- macroalgae %>%
  dplyr::group_by(year, site, habitat,site_habitat) %>%
  dplyr::summarize(across(acanthophora_spicifera:valonia_ventricosa, \(x) mean(x, na.rm = TRUE)))

site_meta_cols <- c("year", "site", "habitat", "site_habitat")
site_meta_df <- site_macro[,which(names(site_macro) %in% meta_cols)]
```

# Figure 1

## setup

```{r cover plot setup}
cover_df <- data.frame(
  year = filtered_data$year,
  habitat = filtered_data$habitat,
  site = filtered_data$site,
  location = filtered_data$location,
  coral = filtered_data$coral,
  cca = filtered_data$crustose_corallines +
    filtered_data$lithophyllum_kotschyanum +
    filtered_data$neogoniolithon_brassica_florida +
    filtered_data$sporolithon_sp,
  macroalgae = rowSums(macro_species_cols),
  turf = filtered_data$algal_turf  + filtered_data$damselfish_turf, 
  sand = filtered_data$sand,
  coral_rubble = filtered_data$coral_rubble,
  cyanophyta = filtered_data$cyanophyta,
  other = filtered_data$ascidian + 
    filtered_data$corallimorpharia +
    filtered_data$heteractis_sp + 
    filtered_data$millepora_platyphylla +
    filtered_data$sarcophyton_sp + filtered_data$shell_debris +
    filtered_data$soft_coral + filtered_data$sponge + filtered_data$tridacna_sp +
    filtered_data$zooxanthellae + filtered_data$no_data + filtered_data$bare_space
)

cover_df$total <- rowSums(cover_df[,-which(names(cover_df) %in% meta_cols)])

cover_means <- cover_df %>%
  group_by(year, habitat) %>%
  summarise(macroalgae = mean(macroalgae),
            coral = mean(coral),
            turf = mean(turf),
            cca = mean(cca),
            sand = mean(sand),
            other = mean(other)
            )

cover_se <- cover_df %>%
  group_by(year, habitat) %>%
  summarise(macroalgae = se(macroalgae),
            coral = se(coral),
            turf = se(turf),
            cca = se(cca),
            sand = se(sand),
            other = se(other)
            )
```

```{r cover long}
cover_long_means <- pivot_longer(
  data = cover_means,
  cols = 3:8,
  names_to = "functional_group",
  values_to = "percent_cover"
)

cover_long_se <- pivot_longer(
  data = cover_se,
  cols = 3:8,
  names_to = "functional_group",
  values_to = "se"
)

cover_long <- left_join(cover_long_means,cover_long_se, by = join_by(year, habitat, functional_group))

#make levels of functional group for plotting
cover_long$functional_group <- factor(cover_long$functional_group, 
                             levels = c("macroalgae", "coral", "turf","cca", "sand", "other"),
                             labels = c("Macroalgae", "Coral", "Turf", "CCA", "Sand", "Other"))
```

```{r make figure1 data}
figure_1_data <- subset(cover_long, functional_group != "Sand" & functional_group != "Other")
```

## plot
```{r figure 1, fig.height = 10, fig.width=14, warning=FALSE}
(figure_1 <- ggplot() +
  geom_point(data = figure_1_data, aes(x = year, y = percent_cover, colour = functional_group), size = 3) +
  geom_errorbar(data = figure_1_data, aes(x = year, ymin = percent_cover - se, ymax = percent_cover + se, colour = functional_group),
                width = 0, size = 1.3) +
   geom_line(data = figure_1_data, aes(x = year, y = percent_cover, colour = functional_group), linewidth = 1) +
  facet_wrap(~ habitat) +
  annotate("segment", x = 2006, xend = 2010, y = 75, yend = 75, linewidth = 2, colour = "grey") +
  scale_colour_manual(values = functional_group_colours) +
  # scale_fill_manual(values = functional_group_colours) +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(cover_long$year)) +
   labs(y = "Percent Cover \n",
        x = "Year") +
  theme_bw() + 
  mytheme)

# ggsave(filename = "output/figure_1.png", figure_1, height = 10, width = 14)
```

## plot after aes changes
```{r Figure 1 v2}
plot_list <- lapply(levels(figure_1_data$habitat), function(hab) {
  ggplot(data = subset(figure_1_data, habitat == hab)) +
    geom_point(aes(x = year, y = percent_cover, colour = functional_group), size = 3) +
    geom_errorbar(aes(x = year, ymin = percent_cover - se, ymax = percent_cover + se, colour = functional_group), 
                  width = 0, size = 1.3) +
    geom_line(aes(x = year, y = percent_cover, colour = functional_group), linewidth = 1) +
    annotate("segment", x = 2006, xend = 2010, y = 75, yend = 75, linewidth = 2, colour = "grey") +
    geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
    geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
    scale_colour_manual(values = functional_group_colours) +
    scale_x_continuous(breaks = unique(figure_1_data$year)) +
    labs(y = "Percent Cover",
         x = "Year",
         title = hab) +  # Add habitat as title
    theme_bw() +
    mytheme
})

figure_1 <- ggarrange(plotlist = plot_list, 
                      ncol = 2, nrow = 2, 
                      labels = c("a.", "b.", "c.", "d."),
                      common.legend = TRUE,
                      legend = "bottom", 
                      font.label = list(size = 20, color = "black", face = "plain"))
# 
# figure_1_caption <- "Figure 1. Benthic coverage of four major space holders in Moorea across four habitats (Fringing, Backreef, Forereef at 10m depth, Forereef  at 17m depth). Annual fluctuations in percent cover of macroalgae, coral, turf, and crustose coralline algae (CCA) are displayed as points for the average cover for each benthic group across sites at each habitat with lines showing standard error. Dotted lines show disturbance events for Cyclone Oli (2010) in black, and the 2019 bleaching event in red, and the gray horizontal line shows crown of thorns (COTs) outbreak which occurred from 2007-2010."
# 
# figure_1_with_caption <- annotate_figure(figure_1,
#                 bottom = text_grob(figure_1_caption, hjust = 0.5, vjust = 1, size = 10))

#ggsave(filename = "output/figure_1v2.png", figure_1, height = 10, width = 14)
```

# Figure 2

## Create Data Frames
```{r create data for figure 2 quad}
alpha_diversity_quad_macro <-
  data.frame(
    year = macroalgae$year,
    location = macroalgae$location,
    shannon = diversity(macroalgae[, -which(names(macroalgae) %in% meta_cols)], "shannon"),
    richness = rowSums(macroalgae[, -which(names(macroalgae) %in% meta_cols)] > 0)
  )

#JG Add Evenness
alpha_diversity_quad_macro$ev <- alpha_diversity_quad_macro$shannon /
  (log(alpha_diversity_quad_macro$richness))

#makes 1 species 0 and 0 species NA for evenness metric
alpha_diversity_quad_macro$evenness <- ifelse(
  alpha_diversity_quad_macro$richness == 0,
  NA,
  # If richness == 0, ev is NA
  ifelse(
    alpha_diversity_quad_macro$richness == 1,
    0,
    # If richness == 1, ev is 0
    alpha_diversity_quad_macro$ev
  )
)  

alpha_diversity_quad_macro <- merge(meta_df, alpha_diversity_quad_macro, by = c("location", "year"))
```

```{r site level richness df}
alpha_diversity_site_macro <-
  data.frame(
    year = site_macro$year,
    site_habitat = site_macro$site_habitat,
    shannon = diversity(site_macro[ , -which(names(site_macro) %in% meta_cols)], "shannon"),
    richness = rowSums(site_macro[ , -which(names(site_macro) %in% meta_cols)] > 0)
  ) %>% 
  left_join(site_meta_df, alpha_diversity_site_macro, by = c("site_habitat", "year"))
```

```{r site level evenness}
#JG Add Evenness
alpha_diversity_site_macro$ev <- alpha_diversity_site_macro$shannon /
  (log(alpha_diversity_site_macro$richness))

#makes 1 species 0 and 0 species NA for evenness metric
alpha_diversity_site_macro$evenness <- ifelse(
  alpha_diversity_site_macro$richness == 0,
  NA,
  # If richness == 0, ev is NA
  ifelse(
    alpha_diversity_site_macro$richness == 1,
    0,
    # If richness == 1, ev is 0
    alpha_diversity_site_macro$ev
  )
)  

alpha_diversity_site_macro <- left_join(site_meta_df, alpha_diversity_site_macro,by = join_by(year, site, habitat, site_habitat))
```

```{r}
macroalgal_cover <- cover_df[,c("year", "habitat", "site", "location", "macroalgae")]

alpha_diversity_quad_macro_with_cover <- left_join(alpha_diversity_quad_macro, macroalgal_cover,by = join_by(location, year, site, habitat)) %>% 
  mutate(macroalgal_prop_cover = macroalgae / 100) %>% 
  select(-macroalgae)
```


## Figure 2a Richness @ Quad by Year

### prep
```{r richness df}
quad_richness_df <- alpha_diversity_quad_macro %>% 
  group_by(year, habitat) %>% 
  summarise(richness_mean = mean(richness),
            richness_se = se(richness))
```


### plot
```{r figure 2a}
(figure_2a <- ggplot() +
  geom_point(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 4) +
  geom_line(data = quad_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
  geom_errorbar(data = quad_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se,
                                             colour = habitat), width = 0, size = 1.3) +
  scale_colour_manual(values = habitat_colours) +
  # scale_fill_manual(values = habitat_colours) +
  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
  annotate("segment", x = 2006, xend = 2010, y = 2, yend = 2, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(quad_richness_df$year)) +
   labs(y = "Macroalgal Richness",
        x = "Year",
        title = "a. Quadrat-level Richness") +
  theme_bw() + 
  mytheme +
   theme(axis.text.x = element_text(size=20),
         axis.title.x = element_text(size=22),
         axis.title.y = element_text(size=22),
      axis.text.y = element_text(size=20),
      plot.title = element_text(size=24),
      legend.text = element_text(size=20),
      legend.title = element_text(size=22)))

# ggsave(filename = "output/figure_2a.png", figure_2a, height = 10, width = 14)
```

## Figure 2b Richness @ Site by Year

### prep

```{r site level richness df}
site_richness_df <- alpha_diversity_site_macro %>% 
  group_by(year, habitat) %>% 
  summarise(richness_mean = mean(richness),
            richness_se = se(richness))

```


### plot
```{r plot figure 2b}
(figure_2b <- ggplot() +
  geom_point(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), size = 4) +
#  geom_jitter(data = alpha_diversity_site_macro, 
#  aes(x = year, y = richness, colour = habitat, alpha = 0.05), size = 2, width = 0.4, height = 0) +
  geom_line(data = site_richness_df, aes(x = year, y = richness_mean, colour = habitat), linewidth = 1) +
  geom_errorbar(data = site_richness_df, aes(x = year, ymin = richness_mean - richness_se, ymax = richness_mean + richness_se, colour = habitat), size = 1.3, width = 0) +
  scale_colour_manual(values = habitat_colours) +
  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
  annotate("segment", x = 2006, xend = 2010, y = 14, yend = 14, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(site_richness_df$year)) +
   labs(y = "",
        x = "Year",
        title = "b. Site-level Richness") +
  mytheme +
   theme(axis.text.x = element_text(size=20),
         axis.title.x = element_text(size=22),
      axis.text.y = element_text(size=20),
      plot.title = element_text(size=24),
      legend.text = element_text(size=20),
      legend.title = element_text(size=22),
      ))

# ggsave(filename = "output/figure_2b.png", figure_2b, height = 10, width = 14)
```

## Figure 2 Full Plot
```{r ggpubr fig2, fig.height = 10, fig.width=20}
fig2_plot_list <- list(figure_2a, figure_2b)
(figure_2 <- ggarrange(plotlist = fig2_plot_list, 
                      ncol = 2, nrow = 1, 
                      common.legend = TRUE,
                      legend = "bottom", 
                      font.label = list(size = 20, color = "black", face = "plain")))

# ggsave(filename = "output/figure_2v2.png", figure_2, height = 10, width = 20)
```


# Figure 3

## set up site-level data for model
```{r site cover by diversity data setup}
# alpha_diversity_site_macro_with_cover <- alpha_diversity_site_macro %>% 
#   dplyr::mutate(percent_cover = rowSums(site_macro[,-which(names(site_macro)  %in% colnames(site_meta_df))])) %>% 
#   mutate(prop_cover = percent_cover / 100) %>% 
#   select(-percent_cover) %>% 
#   mutate(cover_trans = sv_trans(prop_cover))

#there is no reasonable reason why the above chunk doesn't work, but it doesn't, so we did it this way
alpha_diversity_site_macro$percent_cover <- rowSums(site_macro[,-which(names(site_macro)  %in% colnames(site_meta_df))]) 

alpha_diversity_site_macro_with_cover <- alpha_diversity_site_macro %>% 
  dplyr::mutate(prop_cover = percent_cover / 100) %>%
  select(-percent_cover)
  
alpha_diversity_site_macro_with_cover$cover_trans <- sv_trans(alpha_diversity_site_macro_with_cover$prop_cover)
```


## models

### quad
```{r figure 3a richness quad model}
alpha_diversity_quad_macro_with_cover$cover_trans <- sv_trans(alpha_diversity_quad_macro_with_cover$macroalgal_prop_cover)

cover_mod <- glmmTMB(cover_trans ~ richness*habitat + (1|site/location) + (1|year), family = beta_family(), data = alpha_diversity_quad_macro_with_cover)
summary(cover_mod)
Anova(cover_mod)
#                     Chisq Df Pr(>Chisq)    
# richness         30916.73  1  < 2.2e-16 ***
# habitat            141.78  3  < 2.2e-16 ***
# richness:habitat   294.54  3  < 2.2e-16 ***
hist(residuals(cover_mod)) # looks fine
plot(residuals(cover_mod) ~ fitted(cover_mod)) # negative trend but it can't be less than 0 so not too concerned. some wave pattern
r.squaredGLMM(cover_mod) # 0.631, 0.650
emtrends(cover_mod, pairwise ~ habitat, var = "richness") # backreef not different from forereef, forereef also not different
```

```{r 3b evenness quad level model}
alpha_diversity_quad_macro_with_cover_evenness_mod <- alpha_diversity_quad_macro_with_cover %>% 
  drop_na(evenness)

evenness_cover_mod_quad <- glmmTMB(cover_trans ~ evenness*habitat + (1|site/location) + (1|year), family = beta_family(), data = alpha_diversity_quad_macro_with_cover_evenness_mod)
summary(evenness_cover_mod_quad)
Anova(evenness_cover_mod_quad)
#                     Chisq Df Pr(>Chisq)    
# richness         30916.73  1  < 2.2e-16 ***
# habitat            141.78  3  < 2.2e-16 ***
# richness:habitat   294.54  3  < 2.2e-16 ***
hist(residuals(evenness_cover_mod_quad)) # slight pos skeq
plot(residuals(evenness_cover_mod_quad) ~ fitted(evenness_cover_mod_quad)) # negative trend but it can't be less than 0 so not too concerned. some wave pattern
r.squaredGLMM(evenness_cover_mod_quad) # 0.1749907, 0.4042826
emtrends(evenness_cover_mod_quad, pairwise ~ habitat, var = "evenness") #backreef is different from everything except outer 17, but just barely. everyone else is same same xoxo

```

### site richness
```{r site richness mod}
site_cover_mod <- glmmTMB(cover_trans ~ richness*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = alpha_diversity_site_macro_with_cover)
summary(site_cover_mod)
Anova(site_cover_mod)
#                    Chisq Df Pr(>Chisq)    
# richness         58.9195  1  1.643e-14 ***
# habitat           6.6247  3    0.08487 .  
# richness:habitat  3.2967  3    0.34811    
hist(residuals(site_cover_mod)) # some positive skew, not the worst thing ever
plot(residuals(site_cover_mod) ~ fitted(site_cover_mod)) # a little heteroscedastic but can't be less than 0
r.squaredGLMM(site_cover_mod) # 0.2212534, 0.5990825
emtrends(site_cover_mod, pairwise ~ habitat, var = "richness") # no difference between the habitats
```

### site evenness 
```{r}
alpha_diversity_site_macro_with_cover_evenness_mod <- alpha_diversity_site_macro_with_cover %>% 
  drop_na(evenness)

evenness_site_cover_mod <- glmmTMB(cover_trans ~ evenness*habitat + (1|site_habitat) + (1|year), family = beta_family(), data = alpha_diversity_site_macro_with_cover_evenness_mod)
summary(evenness_site_cover_mod)
Anova(evenness_site_cover_mod)
#                    Chisq Df Pr(>Chisq)    
# richness         58.9195  1  1.643e-14 ***
# habitat           6.6247  3    0.08487 .  
# richness:habitat  3.2967  3    0.34811    
hist(residuals(evenness_site_cover_mod)) # some positive skew, not the worst thing ever
plot(residuals(evenness_site_cover_mod) ~ fitted(evenness_site_cover_mod)) # a little heteroscedastic but can't be less than 0
r.squaredGLMM(evenness_site_cover_mod) #0.1206569, 0.5852773
emtrends(evenness_site_cover_mod, pairwise ~ habitat, var = "evenness") # only the outers are different from each other
```

## figure 3a plot setup
```{r figure 3a setup}
diversity_ranges_quad <- alpha_diversity_quad_macro_with_cover %>%
  group_by(habitat) %>%
  summarise(min_richness = min(richness),
            max_richness = max(richness),
            min_cover= min(cover_trans),
            max_cover = max(cover_trans))

cover_emm <- emmip(cover_mod,
                        habitat ~ richness,
                        at = list(richness = seq(0,10,1)), plotit = F, CIs = T)

filtered_cover_effects <- cover_emm %>%
  left_join(diversity_ranges_quad, by = c("habitat")) %>%
  filter(richness >= min_richness & richness <= max_richness)

# alpha_diversity_quad_macro_with_cover$habitat <- factor(alpha_diversity_quad_macro_with_cover$habitat, 
#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
# 
# 
# filtered_cover_effects$habitat <- factor(filtered_cover_effects$habitat, 
#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
```

## figure 3a quad level richness versus cover

```{r figure 3a plot}
(figure_3a <- 
  ggplot() +
  geom_point(data = alpha_diversity_quad_macro_with_cover,
             aes(x = richness, y = cover_trans, colour = habitat), size = 1, alpha = 0.005) + 
  geom_line(data = filtered_cover_effects, aes(x = richness, y = inv_logit(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_cover_effects, aes(x = richness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Cover\n", x = "Macroalgal Richness", title = "3a.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )
# ggsave(filename = "output/figure_3a.png", figure_3a, height = 10, width = 14)
```

## Figure 3B quad level evenness versus cover
```{r figure 3b setup}
diversity_ranges_quad_evenness <- alpha_diversity_quad_macro_with_cover_evenness_mod %>%
  group_by(habitat) %>%
  summarise(min_evenness = min(evenness),
            max_evenness = max(evenness),
            min_cover= min(cover_trans),
            max_cover = max(cover_trans))

cover_emm_evenness <- emmip(evenness_cover_mod_quad,
                        habitat ~ evenness,
                        at = list(evenness = seq(0,10,1)), plotit = F, CIs = T)

filtered_cover_effects_evenness <- cover_emm_evenness %>%
  left_join(diversity_ranges_quad_evenness, by = c("habitat")) %>%
  filter(evenness >= min_evenness & evenness <= max_evenness)

# alpha_diversity_quad_macro_with_cover_evenness_mod$habitat <- factor(alpha_diversity_quad_macro_with_cover_evenness_mod$habitat, 
#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
# 
# 
# filtered_cover_effects_evenness$habitat <- factor(filtered_cover_effects_evenness$habitat, 
#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
```

```{r}
(figure_3b <- 
  ggplot() +
  geom_point(data = alpha_diversity_quad_macro_with_cover_evenness_mod,
             aes(x = evenness, y = cover_trans, colour = habitat), size = 1, alpha = 0.005) + 
  geom_line(data = filtered_cover_effects_evenness, aes(x = evenness, y = inv_logit(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_cover_effects_evenness, aes(x = evenness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Cover\n", x = "Macroalgal evenness", title = "3b.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

#ggsave(filename = "output/figure_3b.png", figure_3b, height = 10, width = 14)
```

## Figure 3c

### synchrony set up
```{r calculate syncrony}
# make er long
macro_long_data <- pivot_longer(
  data = macroalgae,
  cols = grep("acanthophora_spicifera", colnames(macroalgae)):grep("valonia_ventricosa", colnames(macroalgae)),
  names_to = "taxa",
  values_to = "percent_cover") %>% 
  mutate(prop_cover = percent_cover/100) %>% 
  select(-percent_cover)

# write_csv(macro_long_data, here("Data", "macro_long_data.csv"))

# CALCULATE SYNCHRONY CODE
lm_synchrony <- codyn::synchrony(df = macro_long_data,
                                   time.var = "year",
                                   species.var = "taxa",
                                   abundance.var = "prop_cover",
                                   metric = "Loreau",
                                   replicate.var = "location")


diversity_stability <- alpha_diversity_quad_macro_with_cover %>%
  dplyr::group_by(location, habitat, site) %>%
  dplyr::summarise(
            richness_mean = mean(richness), 
            shannon_mean = mean(shannon), 
            cover_mean = mean(macroalgal_prop_cover),
            evenness_mean = mean(evenness, na.rm = T),
            cover_cv = CV(macroalgal_prop_cover),
            cover_stability = 1/CV(macroalgal_prop_cover)) 

diversity_stability_synchrony <- left_join(diversity_stability, lm_synchrony, by = join_by(location))
diversity_stability_synchrony$synchrony_trans <- sv_trans(diversity_stability_synchrony$synchrony)
# diversity_stability_synchrony <- na.omit(diversity_stability_synchrony) # remove quads that never had algae --> NaN calcs


dss_ranges<- diversity_stability_synchrony %>%
  group_by(habitat) %>%
  summarise(min_richness = min(richness_mean),
            max_richness = max(richness_mean),
            min_evenness = min(richness_mean, na.rm = TRUE),
            max_evenness = max(richness_mean, na.rm = TRUE),
            min_stability= min(cover_stability, na.rm = TRUE),
            max_stability = max(cover_stability, na.rm = TRUE),
            min_synchrony = min(synchrony_trans, na.rm = TRUE),
            max_synchrony = max(synchrony_trans, na.rm = TRUE))
```

### synchrony and cover model

```{r synchrony cover mod}
synch_cov_mod <- glmmTMB(cover_mean ~ synchrony*habitat + (1|site), family = beta_family(), data = diversity_stability_synchrony)
summary(synch_cov_mod)
Anova(synch_cov_mod)
#                     Chisq Df Pr(>Chisq)    
# synchrony         209.545  1  < 2.2e-16 ***
# habitat            52.512  3  2.330e-11 ***
# synchrony:habitat  54.632  3  8.227e-12 ***
hist(residuals(synch_cov_mod)) # slight skew
plot(residuals(synch_cov_mod) ~ fitted(synch_cov_mod)) # same issues as always

r.squaredGLMM(synch_cov_mod) # error??
emtrends(synch_cov_mod, pairwise ~ habitat, var = "synchrony") # forereef 17 distinct from all other habitats
```

```{r synch cov plot setup}
synch_cov_emm <- emmip(synch_cov_mod,
                        habitat ~ synchrony,
                        at = list(synchrony = seq(0,1,0.01)), plotit = F, CIs = T)

filtered_synch_cov_effects <- synch_cov_emm %>%
  left_join(dss_ranges, by = c("habitat")) %>%
  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
```

### plot

```{r synchrony cover plot}
(figure_3c <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony,
             aes(x = synchrony, y = cover_mean, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_synch_cov_effects, aes(x = synchrony, y = exp(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_synch_cov_effects, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(x = "Synchrony", y = "Cover\n", title = "3c.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18),
        plot.title = element_text(size=18),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )


# ggsave(filename = "output/figure_3c.png", figure_3c, height = 10, width = 14)
```


### Synchrony setup site level
```{r calculate syncrony site_level}
# make er long
macro_long_data_site <- pivot_longer(
  data = site_macro,
  cols = grep("acanthophora_spicifera", colnames(site_macro)):grep("valonia_ventricosa", colnames(site_macro)),
  names_to = "taxa",
  values_to = "percent_cover") %>% 
  mutate(prop_cover = percent_cover/100) %>% 
  select(-percent_cover)

# write_csv(macro_long_data, here("Data", "macro_long_data.csv"))

# CALCULATE SYNCHRONY CODE
lm_synchrony_site <- codyn::synchrony(df = macro_long_data_site,
                                   time.var = "year",
                                   species.var = "taxa",
                                   abundance.var = "prop_cover",
                                   metric = "Loreau",
                                   replicate.var = "site_habitat")


diversity_stability_site <- alpha_diversity_site_macro_with_cover %>%
  dplyr::group_by(site_habitat, habitat, site) %>%
  dplyr::summarise(
            richness_mean = mean(richness), 
            shannon_mean = mean(shannon), 
            cover_mean = mean(prop_cover),
            evenness_mean = mean(evenness),
            cover_cv = CV(prop_cover),
            cover_stability = 1/CV(prop_cover)) 

diversity_stability_synchrony_site <- left_join(diversity_stability_site, lm_synchrony_site, by = join_by(site_habitat))
diversity_stability_synchrony_site$synchrony_trans <- sv_trans(diversity_stability_synchrony_site$synchrony)
# diversity_stability_synchrony_site <- na.omit(diversity_stability_synchrony_site) # remove quads that never had algae --> NaN calcs
# write_csv(diversity_stability_synchrony_site, here("data", "diversity_stability_synchrony_site.csv"))

dss_ranges_site <- diversity_stability_synchrony_site %>%
  group_by(habitat) %>%
  summarise(min_richness = min(richness_mean),
            max_richness = max(richness_mean),
            min_evenness = min(evenness_mean, na.rm = T),
            max_evenness = max(evenness_mean, na.rm = T),
            min_stability= min(cover_stability),
            max_stability = max(cover_stability),
            min_synchrony = min(synchrony_trans),
            max_synchrony = max(synchrony_trans))
```

## Figure 3E site level richness versus cover

### setup
```{r}
diversity_ranges_site <- alpha_diversity_site_macro_with_cover %>%
  group_by(habitat) %>%
  summarise(min_richness = min(richness),
            max_richness = max(richness),
            min_cover= min(cover_trans),
            max_cover = max(cover_trans))

site_cover_emm <- emmip(site_cover_mod,
                        habitat ~ richness,
                        at = list(richness = seq(0,17,1)), plotit = F, CIs = T)

filtered_site_cover_effects <- site_cover_emm %>%
  left_join(diversity_ranges_site, by = c("habitat")) %>%
  filter(richness >= min_richness & richness <= max_richness)

# alpha_diversity_site_macro_with_cover$habitat <- factor(alpha_diversity_site_macro_with_cover$habitat, 
#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
# 
# 
# filtered_site_cover_effects$habitat <- factor(filtered_site_cover_effects$habitat, 
#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
```

### plot

```{r}
(figure_3d <- 
  ggplot() +
  geom_point(data = alpha_diversity_site_macro_with_cover,
             aes(x = richness, y = cover_trans, colour = habitat), size = 1, alpha = 0.3) + 
  geom_line(data = filtered_site_cover_effects, aes(x = richness, y = inv_logit(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_site_cover_effects, aes(x = richness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Cover\n", x = "Macroalgal Richness", title = "3d.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        plot.title = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )
# ggsave(filename = "output/figure_3d.png", figure_3d, height = 10, width = 14)
```

## figure 3f

### model

```{r synchrony cover mod}
synch_cov_mod_site <- glmmTMB(cover_mean ~ synchrony*habitat, family = beta_family(), data = diversity_stability_synchrony_site)
summary(synch_cov_mod_site)
Anova(synch_cov_mod_site)
#                    Chisq Df Pr(>Chisq)  
# synchrony         2.4184  1    0.11992  
# habitat           7.8002  3    0.05033 .
# synchrony:habitat 8.8866  3    0.03084 *
hist(residuals(synch_cov_mod_site)) # slight skew
plot(residuals(synch_cov_mod_site) ~ fitted(synch_cov_mod_site)) # looks fine
emtrends(synch_cov_mod_site, pairwise ~ habitat, var = "synchrony") # forereef 17 different from backreef
```

```{r synch cov site plot setup}
synch_cov_emm_site<- emmip(synch_cov_mod_site,
                        habitat ~ synchrony,
                        at = list(synchrony = seq(0,1,0.01)), plotit = F, CIs = T)

filtered_synch_cov_effects_site <- synch_cov_emm_site %>%
  left_join(dss_ranges_site, by = c("habitat")) %>%
  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
```

### plot figure 3f

```{r synchrony cover plot site}
(figure_3f <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony_site,
             aes(x = synchrony, y = cover_mean, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_synch_cov_effects_site, aes(x = synchrony, y = exp(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_synch_cov_effects_site, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(x = "Synchrony", y = "Cover\n", title = "3f.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        plot.title = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )


# ggsave(filename = "output/figure_3f.png", figure_3f, height = 10, width = 14)
```

## Figure 3f site level evenness versus cover

### model

```{r evenness cover model site level}
evenness_cover_mod_site <- glmmTMB(cover_trans ~ evenness*habitat + (1|site) + (1|year), family = beta_family(), data = alpha_diversity_site_macro_with_cover_evenness_mod)
summary(evenness_cover_mod_site)
hist(residuals(evenness_cover_mod_site)) # little wonky but not the worst
plot(residuals(evenness_cover_mod_site) ~ fitted(evenness_cover_mod_site))
Anova(evenness_cover_mod_site)
#                   Chisq Df Pr(>Chisq)    
# evenness         14.395  1  0.0001482 ***
# habitat          45.636  3  6.777e-10 ***
# evenness:habitat 16.126  3  0.0010687 ** 
r.squaredGLMM(evenness_cover_mod_site) # 0.166, 0.258

emtrends(evenness_cover_mod_site, pairwise ~ habitat, var = "evenness") # fringing different from backreef and outer10
```

### plot

### setup

```{r figure 3f evenness site plot setup}
diversity_ranges_site_evenness <- alpha_diversity_site_macro_with_cover_evenness_mod %>%
  group_by(habitat) %>%
  summarise(min_evenness = min(evenness),
            max_evenness = max(evenness),
            min_cover= min(cover_trans),
            max_cover = max(cover_trans))

cover_emm_evenness_site <- emmip(evenness_cover_mod_site,
                        habitat ~ evenness,
                        at = list(evenness = seq(0,1,0.01)), plotit = F, CIs = T)

filtered_cover_effects_evenness_site <- cover_emm_evenness_site %>%
  left_join(diversity_ranges_site_evenness, by = c("habitat")) %>%
  filter(evenness >= min_evenness & evenness <= max_evenness)

# alpha_diversity_site_macro_with_cover_evenness_mod$habitat <- factor(alpha_diversity_site_macro_with_cover_evenness_mod$habitat, 
#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
# 
# 
# 
```

### plot

```{r evenness cover plot site }
(figure_3f <- 
  ggplot() +
  geom_point(data = alpha_diversity_site_macro_with_cover_evenness_mod,
             aes(x = evenness, y = cover_trans, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_cover_effects_evenness_site, aes(x = evenness, y = inv_logit(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_cover_effects_evenness_site, aes(x = evenness, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Cover\n", x = "Macroalgal evenness", title = "3f. Site Level Evenness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

# ggsave(filename = "output/figure_3f.png", figure_3f, height = 10, width = 14)
```


# Figure 4


## Figure 4a 

###model - richness stability

```{r richness synchrony mod}
rich_stab_mod <- glmmTMB(cover_stability~ richness_mean*habitat + (1|site), family = Gamma(link = "log"), data = diversity_stability_synchrony)
summary(rich_stab_mod)
Anova(rich_stab_mod)
#                          Chisq Df Pr(>Chisq)    
# richness_mean         2721.438  1  < 2.2e-16 ***
# habitat                325.876  3  < 2.2e-16 ***
# richness_mean:habitat   33.926  3  2.054e-07 ***
hist(residuals(rich_stab_mod)) # looks good
plot(residuals(rich_stab_mod) ~ fitted(rich_stab_mod)) # heteroskedastic - bring to group
r.squaredGLMM(rich_stab_mod) # 0.803, 0.808
emtrends(rich_stab_mod, pairwise ~ habitat, var = "richness_mean") # fringing not diff from outer17, backreef not different from outer10
```

```{r figure 4a setup}
rich_stab_emm <- emmip(rich_stab_mod,
                        habitat ~ richness_mean,
                        at = list(richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)

filtered_rich_stab_effects <- rich_stab_emm %>%
  left_join(dss_ranges, by = c("habitat")) %>%
  filter(richness_mean >= min_richness & richness_mean <= max_richness)
```


### plot figure 4a

```{r figure 4a}
(figure_4a <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony,
             aes(x = richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_rich_stab_effects, aes(x = richness_mean, y = exp(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_rich_stab_effects, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Stability\n", x = "Taxonomic Richness", title = "4a.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18),
        plot.title = element_text(size=18),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

# ggsave(filename = "output/figure_4a.png", figure_4a, height = 10, width = 14)
```

#### no habitat

```{r richness stability no hab}
# rich_stab_mod_2 <- glmmTMB(cover_stability~ richness_mean + (1|site), family = Gamma(link = "log"), data = # diversity_stability_synchrony)
# summary(rich_stab_mod_2)
# hist(residuals(rich_stab_mod_2)) # a chunkier version
# plot(residuals(rich_stab_mod_2) ~ fitted(rich_stab_mod_2)) # heteroskedastic - bring to group
# r.squaredGLMM(rich_stab_mod_2) # 0.750, 0.757


```

```{r 4a no hab setup}
# rich_stab_emm_2 <-as.data.frame(emmeans(rich_stab_mod_2, ~richness_mean,
#                                      at = list(richness_mean = seq(min(diversity_stability_synchrony$richness_mean),
#                                                             max(diversity_stability_synchrony$richness_mean),
#                                                             0.01)))) 
# 
```

#### no habitat plot

```{r figure 4a no habitat}
# (figure_4a_no_hab <- 
#   ggplot() +
#   geom_point(data = diversity_stability_synchrony,
#              aes(x = richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
#   geom_line(data = filtered_rich_stab_effects, aes(x = richness_mean, y = exp(yvar), colour = habitat)) +
#   geom_ribbon(data = filtered_rich_stab_effects, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
#               alpha = 0.3) +
#   geom_line(data = rich_stab_emm_2, aes(x = richness_mean, y = exp(emmean)), size = 2, colour = "black") +
#   geom_ribbon(data = rich_stab_emm_2, aes(x = richness_mean, ymin = exp(asymp.LCL), ymax = exp(asymp.UCL)), fill = "black",
#               alpha = 0.3) +
#   scale_colour_manual(values = habitat_colours) +
#   scale_fill_manual(values = habitat_colours) +
#   labs(y = "Stability\n", x = "Macroalgal Richness") +
#   theme_bw() + 
#   theme(axis.title.x = element_text(size=18), 
#         axis.title.y = element_text(size=18), 
#         panel.background = element_blank(), 
#         panel.grid.major = element_blank(),  #remove major-grid labels
#         panel.grid.minor = element_blank(),  #remove minor-grid labels
#         plot.background = element_rect(fill = "white"),
#         legend.text = element_text(size = rel(1)), # increase legend text
#         legend.title = element_blank(),
#         legend.position = 'bottom', # move legend to bottom
#         legend.box = 'horizontal', # make legend horizontal
#         legend.box.background = element_rect(colour = "black"))  # add box around legend
#          )
# 
# ggsave(filename = "output/figure_4a_with_no_hab.png", figure_4a_no_hab, height = 10, width = 14)
```

## figure 4c 
### model -  synchrony staability

```{r synchrony staability mod}
synch_stab_mod <- glmmTMB(cover_stability ~ synchrony*habitat + (1|site), family = Gamma("log"), data = diversity_stability_synchrony)
summary(synch_stab_mod)
Anova(synch_stab_mod)
#                     Chisq Df Pr(>Chisq)    
# synchrony         1308.33  1  < 2.2e-16 ***
# habitat            103.13  3  < 2.2e-16 ***
# synchrony:habitat  327.97  3  < 2.2e-16 ***
hist(residuals(synch_stab_mod)) # looks good
plot(residuals(synch_stab_mod) ~ fitted(synch_stab_mod)) # same issue as always
r.squaredGLMM(synch_stab_mod) # 0.63, 0.71
emtrends(synch_stab_mod, pairwise ~ habitat, var = "synchrony") # fringing and backreef not different
```

```{r figure 4c setup}
synch_stab_emm <- emmip(synch_stab_mod,
                        habitat ~ synchrony,
                        at = list(synchrony = seq(0,1,0.01)), plotit = F, CIs = T)

filtered_synch_stab_effects <- synch_stab_emm %>%
  left_join(dss_ranges, by = c("habitat")) %>%
  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
```

### plot figure 4c

```{r figure 4c}
(figure_4c <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony,
             aes(x = synchrony, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_synch_stab_effects, aes(x = synchrony, y = exp(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_synch_stab_effects, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(x = "Synchrony", y = "Stability\n", title = "4c.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18),
        plot.title = element_text(size=18),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )
# ggsave(filename = "output/figure_4c.png", figure_4c, height = 10, width = 14)
```

## figure 4d 

###model - richness stability

```{r richness synchrony mod}
rich_stab_mod_site <- glmmTMB(cover_stability~ richness_mean*habitat, family = Gamma(link = "log"), data = diversity_stability_synchrony_site)
summary(rich_stab_mod_site)
Anova(rich_stab_mod_site)

#                        Chisq Df Pr(>Chisq)    
# richness_mean         17.645  1  2.662e-05 ***
# habitat               32.229  3  4.682e-07 ***
# richness_mean:habitat  8.075  3    0.04449 *  
hist(residuals(rich_stab_mod_site)) # blocky but fine
plot(residuals(rich_stab_mod_site) ~ fitted(rich_stab_mod_site)) # heteroskedastic - bring to group
r.squaredGLMM(rich_stab_mod_site) # 0.671
emtrends(rich_stab_mod_site, pairwise ~ habitat, var = "richness_mean") # no differences across habitats
```

```{r figure 4d setup}
rich_stab_emm_site <- emmip(rich_stab_mod_site,
                        habitat ~ richness_mean,
                        at = list(richness_mean = seq(0,10,0.01)), plotit = F, CIs = T)

filtered_rich_stab_effects_site <- rich_stab_emm_site %>%
  left_join(dss_ranges_site, by = c("habitat")) %>%
  filter(richness_mean >= min_richness & richness_mean <= max_richness)
```


### plot figure 4d

```{r figure 4d}
(figure_4d <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony_site,
             aes(x = richness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_rich_stab_effects_site, aes(x = richness_mean, y = exp(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_rich_stab_effects_site, aes(x = richness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Stability\n", x = "Macroalgal Richness", title = "4d.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        plot.title = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

# ggsave(filename = "output/figure_4d.png", figure_4d, height = 10, width = 14)
```


## figure 4f

### model
```{r synchrony staability mod site}
synch_stab_mod_site <- glmmTMB(cover_stability ~ synchrony*habitat, family = Gamma("log"), data = diversity_stability_synchrony_site)
summary(synch_stab_mod_site)
Anova(synch_stab_mod_site)
#                     Chisq Df Pr(>Chisq)    
# synchrony         22.0686  1  2.631e-06 ***
# habitat            9.3333  3    0.02517 *  
# synchrony:habitat 10.5561  3    0.01439 *  
hist(residuals(synch_stab_mod_site)) # almost uniform
plot(residuals(synch_stab_mod_site) ~ fitted(synch_stab_mod_site)) # same issue as always
r.squaredGLMM(synch_stab_mod_site) # 0.678
emtrends(synch_stab_mod_site, pairwise ~ habitat, var = "synchrony") # backreef is different from forereef 17, fringing is a p = 0.05
```

```{r figure 4f setup}
synch_stab_emm_site <- emmip(synch_stab_mod_site,
                        habitat ~ synchrony,
                        at = list(synchrony = seq(0,1,0.01)), plotit = F, CIs = T)

filtered_synch_stab_effects_site <- synch_stab_emm_site %>%
  left_join(dss_ranges_site, by = c("habitat")) %>%
  filter(synchrony >= min_synchrony & synchrony <= max_synchrony)
```

### plot figure 4f

```{r figure 4f}
(figure_4f <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony_site,
             aes(x = synchrony, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_synch_stab_effects_site, aes(x = synchrony, y = exp(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_synch_stab_effects_site, aes(x = synchrony, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(x = "Synchrony", y = "Stability\n", main = "4f.") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18),
        plot.title = element_text(size=18),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )
# ggsave(filename = "output/figure_4f.png", figure_4f, height = 10, width = 14)
```


# evenness stability site

## mod

```{r evenness synchrony mod}
even_stab_mod_site <- glmmTMB(cover_stability~ evenness_mean*habitat, family = Gamma(link = "log"), data = diversity_stability_synchrony_site)
summary(even_stab_mod_site)
Anova(even_stab_mod_site)

#                         Chisq Df Pr(>Chisq)   
# evenness_mean          0.1035  1   0.747623   
# habitat               11.8743  3   0.007827 **
# evenness_mean:habitat  9.6254  3   0.022034 *   
hist(residuals(even_stab_mod_site)) # blocky with some pos skew
plot(residuals(even_stab_mod_site) ~ fitted(even_stab_mod_site)) # heteroskedastic 
r.squaredGLMM(even_stab_mod_site) # 0.505
emtrends(even_stab_mod_site, pairwise ~ habitat, var = "evenness_mean") # no differences across habitats
```

```{r figure 4d setup}
even_stab_emm_site <- emmip(even_stab_mod_site,
                        habitat ~ evenness_mean,
                        at = list(evenness_mean = seq(0,1,0.01)), plotit = F, CIs = T)
#make levels for habitat for plotting


filtered_even_stab_effects_site <- even_stab_emm_site %>%
  left_join(dss_ranges_site, by = c("habitat")) %>%
  filter(evenness_mean >= min_evenness & evenness_mean <= max_evenness)

#make levels for habitat for plotting
# diversity_stability_synchrony_site$habitat <- factor(diversity_stability_synchrony_site$habitat, 
#                              levels = c("fringing", "backreef", "outer_10","outer_17"),
#                              labels = c("Fringing", "Backreef", "Forereef 10m", "Forereef 17m"))
```


## plot figure 4d

```{r figure 4d}
(figure_4_evenness <- 
  ggplot() +
  geom_point(data = diversity_stability_synchrony_site,
             aes(x = evenness_mean, y = cover_stability, colour = habitat), size = 1, alpha = 0.5) + 
  geom_line(data = filtered_even_stab_effects_site, aes(x = evenness_mean, y = exp(yvar), colour = habitat)) +
  geom_ribbon(data = filtered_even_stab_effects_site, aes(x = evenness_mean, ymin = exp(LCL), ymax = exp(UCL), fill = habitat),
              alpha = 0.3) +
  scale_colour_manual(values = habitat_colours) +
  scale_fill_manual(values = habitat_colours) +
  labs(y = "Stability\n", x = "Macroalgal evenness", title = "4f. Evenness site") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.title.y = element_text(size=18), 
        plot.title = element_text(size=18), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
         )

ggsave(filename = "output/figure_4_evenness_site.png", figure_4_evenness, height = 10, width = 14)
```


## Figure 4 Full Plot
```{r ggpubr fig4}
fig4_plot_list <- list(figure_4a, figure_4c, figure_4d, figure_4f)
figure_4 <- ggarrange(plotlist = fig4_plot_list, 
                      ncol = 2, nrow = 2, 
                      labels = c("a.", "b."),
                      common.legend = TRUE,
                      legend = "bottom", 
                      font.label = list(size = 20, color = "black", face = "plain"))

#ggsave(filename = "output/figure_4.png", figure_4, height = 10, width = 20)
```


# SUPPLEMENTAL FIGURE CODE:

## 2c Evenness @ Quad by Year
### prep
```{r calculate evennes}
#dear callie, why is this here?
site_richness_df <- alpha_diversity_site_macro %>% 
  group_by(year, habitat) %>% 
  summarise(richness_mean = mean(richness),
            richness_se = se(richness))

quad_evenness_df <- alpha_diversity_quad_macro_with_cover %>% 
  drop_na(evenness) %>%
  group_by(year, habitat) %>% 
  summarise(evenness_mean = mean(evenness),
            evenness_se = se(evenness))
```

### plot
```{r}
(figure_2c <- ggplot() +
  geom_point(data = quad_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), size = 3) +
  geom_line(data = quad_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), linewidth = 1) +
  geom_errorbar(data = quad_evenness_df, aes(x = year, ymin = evenness_mean - evenness_se, ymax = evenness_mean + evenness_se,
                                             colour = habitat), width = 0, size = 1.3) +
  scale_colour_manual(values = habitat_colours) +
  # scale_fill_manual(values = habitat_colours) +
  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
  annotate("segment", x = 2006, xend = 2010, y = .7, yend = .7, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(quad_evenness_df$year)) +
   labs(y = "Macroalgal evenness \n",
        x = "Year",
        title = "c. Quadrat-level evenness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size=18), 
        plot.title = element_text(size=18),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

# ggsave(filename = "output/figure_2c.png", figure_2c, height = 10, width = 14)
```

## 2d Evenness @ Site by Year
### prep
```{r set up evenness}
site_evenness_df <- alpha_diversity_site_macro %>% 
  drop_na(evenness) %>%
  group_by(year, habitat) %>% 
  summarise(evenness_mean = mean(evenness),
            evenness_se = se(evenness))
```

### plot
```{r plot evenness site}
(figure_2d <- ggplot() +
  geom_point(data = site_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), size = 3) +
  geom_line(data = site_evenness_df, aes(x = year, y = evenness_mean, colour = habitat), linewidth = 1) +
  geom_errorbar(data = site_evenness_df, aes(x = year, ymin = evenness_mean - evenness_se, ymax = evenness_mean + evenness_se, colour = habitat), size = 1.3, width = 0) +
  scale_colour_manual(values = habitat_colours) +
  #geom_segment(aes(y = 75, yend = 75, x = 2006, xend = 2010), linewidth = 2, colour = "grey") +
  annotate("segment", x = 2006, xend = 2010, y = 1, yend = 1, linewidth = 2, colour = "grey") +
  geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
  geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   scale_x_continuous(breaks = unique(site_evenness_df$year)) +
   labs(y = "Macroalgal evenness \n",
        x = "Year",
        title = "d. Site-level evenness") +
  theme_bw() + 
  theme(axis.title.x = element_text(size=18), 
        axis.text.x = element_text(angle = 45, hjust = 1),
        axis.title.y = element_text(size=18), 
        plot.title = element_text(size=18),
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.text = element_text(size = rel(1)), # increase legend text
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black")) )

# ggsave(filename = "output/figure_2d.png", figure_2d, height = 10, width = 14)
```
