---
title: "SEM with quad-level algae data"
output: html_document
date: "2025-01-17"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# clear workspace
rm(list = ls())


require(tidyverse)
require(here)
require(piecewiseSEM)
require(lme4) # glmer() function
require(performance)
```

# Read in data

```{r}
# functional richness from Noam
functrich <- read_csv(here("data/full_plot_level_dss.csv"))

# quad-level data summarized with one point for all time
algae_1 <- read_csv(here("data/diversity_stability_synchrony.csv"))
```

These two guys have different dimensions. See where they differ:

```{r}
functrich[!(functrich$location %in% algae_1$location),] 
```

Noam says we can just inner join/exclude these--just used to calculate averages

```{r}
functrich %>% 
  select(location, habitat, site, functional_richness_mean) %>% 
  inner_join(algae_1) -> algae
```


# Scale data

Gamma families are currently not supported and we're not able to get glmmTMB to work for the time being.

```{r}
algae %>% 
  mutate(
    richness_mean_scale = scale(algae$richness_mean)[,1],
    functional_richness_mean_scale = scale(algae$functional_richness_mean)[,1],
    cover_mean_scale = scale(algae$cover_mean)[,1],
    cover_stability_scale = scale(algae$cover_stability)[,1],
    synchrony_scale = scale(algae$synchrony)[,1]
  ) -> algae_scaled
```


# Make an overall SEM with functional richness

First, fit individual models

```{r}
# effect of richness on synchrony
m.sprich.synch <- glmer(synchrony_scale ~ richness_mean_scale + functional_richness_mean_scale + (1|site) + (1|habitat),
                        data = algae_scaled)
  performance::check_model(m.sprich.synch)
  
# effect of taxonomic richness on functional richness
m.sprich.funrich <- glmer(functional_richness_mean_scale ~ richness_mean_scale + (1|site) + (1|habitat),
                        data = algae_scaled)
  performance::check_model(m.sprich.funrich)

# effect of synchrony on stability
m.synch.stab.cov <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale + functional_richness_mean_scale + (1|site) + (1|habitat),
                      data = algae_scaled)  
  performance::check_model(m.synch.stab.cov)
  
# effect of richness on cover
m.sprich.cov <- glmer(cover_mean_scale ~ richness_mean_scale + (1|site) + (1|habitat),
                      data = algae_scaled)
  performance::check_model(m.sprich.cov)
```

Try it with piecewiseSEM

```{r}
# Build SEM
sem <- psem(
  m.sprich.synch,
  m.sprich.funrich,
  m.synch.stab.cov,
  m.sprich.cov,
  data = as.data.frame(algae_scaled)
)

summary(sem)
```

Plot

```{r}
plot(sem)
```


# Make an overall SEM withOUT functional richness

First, fit individual models

```{r}
# effect of richness on synchrony
m.sprich.synch.noFR <- glmer(synchrony_scale ~ richness_mean_scale  + (1|site) + (1|habitat),
                        data = algae_scaled)
  performance::check_model(m.sprich.synch.noFR)

# effect of synchrony on stability
m.synch.stab.cov.noFR <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale + (1|site) + (1|habitat),
                      data = algae_scaled)  
  performance::check_model(m.synch.stab.cov.noFR)
  
# effect of richness on cover
m.sprich.cov.noFR <- glmer(cover_mean_scale ~ richness_mean_scale + (1|site) + (1|habitat),
                      data = algae_scaled)
  performance::check_model(m.sprich.cov.noFR)
```

Try it with piecewiseSEM

```{r}
# Build SEM
sem.noFR <- psem(
  m.sprich.synch,
  m.synch.stab.cov,
  m.sprich.cov,
  data = as.data.frame(algae_scaled)
)

summary(sem.noFR)
```

Plot

```{r}
plot(sem.noFR)
```

# Frining reef: bracketed 

This is trying to make these models without rescaling within the fringing reef to see if it matters (i.e., we'll just subset from the original scaled dataframe)

```{r}
# effect of richness on synchrony
m.sprich.synch.FR <- glmer(synchrony_scale ~ richness_mean_scale + functional_richness_mean_scale + (1|site),
                        data = algae_scaled[algae_scaled$habitat == "Fringing", ])
  performance::check_model(m.sprich.synch.FR)
  
# effect of taxonomic richness on functional richness
m.sprich.funrich.FR <- glmer(functional_richness_mean_scale ~ richness_mean_scale + (1|site),
                        data = algae_scaled[algae_scaled$habitat == "Fringing", ])
  performance::check_model(m.sprich.funrich.FR)

# effect of synchrony on stability
m.synch.stab.cov.FR <- glmer(cover_stability_scale ~ synchrony_scale + functional_richness_mean_scale +  cover_mean_scale + richness_mean_scale + (1|site),
                      data = algae_scaled[algae_scaled$habitat == "Fringing", ])  
  performance::check_model(m.synch.stab.cov.FR)
  
# effect of richness on cover
m.sprich.cov.FR <- glmer(cover_mean_scale ~ richness_mean_scale + (1|site),
                      data = algae_scaled[algae_scaled$habitat == "Fringing", ])
  performance::check_model(m.sprich.cov.FR)
```

Try it with piecewiseSEM

```{r}
# Build SEM
sem_fringe_brac <- psem(
  m.sprich.synch.FR,
  m.sprich.funrich.FR,
  m.synch.stab.cov.FR,
  m.sprich.cov.FR,
  data = as.data.frame(algae_scaled[algae_scaled$habitat == "Fringing", ])
)

summary(sem_fringe_brac)
```

Plot

```{r}
plot(sem_fringe_brac)
```

# Backreef: bracketed 

This is trying to make these models without rescaling within the fringing reef to see if it matters (i.e., we'll just subset from the original scaled dataframe)

```{r}
# effect of richness on synchrony
m.sprich.synch.BA <- glmer(synchrony_scale ~ richness_mean_scale + functional_richness_mean_scale + (1|site),
                        data = algae_scaled[algae_scaled$habitat == "Backreef", ])
  performance::check_model(m.sprich.synch.BA)
  
# effect of taxonomic richness on functional richness
m.sprich.funrich.BA <- glmer(functional_richness_mean_scale ~ richness_mean_scale + (1|site),
                        data = algae_scaled[algae_scaled$habitat == "Backreef", ])
  performance::check_model(m.sprich.funrich.BA)


# effect of synchrony on stability
m.synch.stab.cov.BA <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + functional_richness_mean_scale + richness_mean_scale + (1|site),
                      data = algae_scaled[algae_scaled$habitat == "Backreef", ])  
  performance::check_model(m.synch.stab.cov.BA)
  
# effect of richness on cover
m.sprich.cov.BA <- glmer(cover_mean_scale ~ richness_mean_scale + (1|site),
                      data = algae_scaled[algae_scaled$habitat == "Backreef", ])
  performance::check_model(m.sprich.cov.BA)
```

Try it with piecewiseSEM

```{r}
# Build SEM
sem_back_brac <- psem(
  m.sprich.synch.BA,
  m.sprich.funrich.BA,
  m.synch.stab.cov.BA,
  m.sprich.cov.BA,
  data = as.data.frame(algae_scaled[algae_scaled$habitat == "Backreef", ])
)

summary(sem_back_brac)
```

Plot

```{r}
plot(sem_back_brac)
```

# Forereef 10: bracketed 

This is trying to make these models without rescaling within the fringing reef to see if it matters (i.e., we'll just subset from the original scaled dataframe)

```{r}
# effect of richness on synchrony
m.sprich.synch.FO10 <- glmer(synchrony_scale ~ richness_mean_scale + functional_richness_mean_scale + (1|site),
                        data = algae_scaled[algae_scaled$habitat == "Forereef 10m", ])
  performance::check_model(m.sprich.synch.FO10)
  
# effect of taxonomic richness on functional richness
m.sprich.funrich.FO10 <- glmer(functional_richness_mean_scale ~ richness_mean_scale + (1|site),
                        data = algae_scaled[algae_scaled$habitat == "Forereef 10m", ])
  performance::check_model(m.sprich.funrich.FO10)

# effect of synchrony on stability
m.synch.stab.cov.FO10 <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale + functional_richness_mean_scale + (1|site),
                      data = algae_scaled[algae_scaled$habitat == "Forereef 10m", ])  
  performance::check_model(m.synch.stab.cov.FO10)
  
# effect of richness on cover
m.sprich.cov.FO10 <- glmer(cover_mean_scale ~ richness_mean_scale + (1|site),
                      data = algae_scaled[algae_scaled$habitat == "Forereef 10m", ])
  performance::check_model(m.sprich.cov.FO10)
```

Try it with piecewiseSEM

```{r}
# Build SEM
sem_for10_brac <- psem(
  m.sprich.synch.FO10,
  m.sprich.funrich.FO10,
  m.synch.stab.cov.FO10,
  m.sprich.cov.FO10,
  data = as.data.frame(algae_scaled[algae_scaled$habitat == "Forereef 10m", ])
)

summary(sem_for10_brac)
```

Plot

```{r}
plot(sem_for10_brac)
```

# Forereef 17: bracketed 

This is trying to make these models without rescaling within the fringing reef to see if it matters (i.e., we'll just subset from the original scaled dataframe)

```{r}
# effect of richness on synchrony
m.sprich.synch.FO17 <- glmer(synchrony_scale ~ richness_mean_scale + functional_richness_mean_scale + (1|site),
                        data = algae_scaled[algae_scaled$habitat == "Forereef 17m", ])
  performance::check_model(m.sprich.synch.FO17)
  
# effect of taxonomic richness on functional richness
m.sprich.funrich.FO17 <- glmer(functional_richness_mean_scale ~ richness_mean_scale + (1|site),
                        data = algae_scaled[algae_scaled$habitat == "Forereef 17m", ])
  performance::check_model(m.sprich.funrich.FO17)

# effect of synchrony on stability
m.synch.stab.cov.FO17 <- glmer(cover_stability_scale ~ synchrony_scale + functional_richness_mean_scale + cover_mean_scale + richness_mean_scale + (1|site),
                      data = algae_scaled[algae_scaled$habitat == "Forereef 17m", ])  
  performance::check_model(m.synch.stab.cov.FO17)
  
# effect of richness on cover
m.sprich.cov.FO17 <- glmer(cover_mean_scale ~ richness_mean_scale + (1|site),
                      data = algae_scaled[algae_scaled$habitat == "Forereef 17m", ])
  performance::check_model(m.sprich.cov.FO17)
```

Try it with piecewiseSEM

```{r}
# Build SEM
sem_for17_brac <- psem(
  m.sprich.synch.FO17,
  m.sprich.funrich.FO17,
  m.synch.stab.cov.FO17,
  m.sprich.cov.FO17,
  data = as.data.frame(algae_scaled[algae_scaled$habitat == "Forereef 17m", ])
)

summary(sem_for17_brac)
```

Plot

```{r}
plot(sem_for17_brac)
```

# Try re-scaling within a habitat: Frining

```{r}
algae %>% 
  filter(habitat == "Fringing") -> fringe
  
fringe %>% 
  mutate(
    richness_mean_scale = scale(fringe$richness_mean)[,1],
    cover_mean_scale = scale(fringe$cover_mean)[,1],
    cover_stability_scale = scale(fringe$cover_stability)[,1],
    synchrony_scale = scale(fringe$synchrony)[,1]
  ) -> algae_scaled_fringe

```

```{r}
# effect of richness on synchrony
m.sprich.synch.FRsc <- glmer(synchrony_scale ~ richness_mean_scale + (1|site),
                        data = algae_scaled_fringe)
  performance::check_model(m.sprich.synch.FRsc)

# effect of synchrony on stability
m.synch.stab.cov.FRsc <- glmer(cover_stability_scale ~ synchrony_scale + cover_mean_scale + richness_mean_scale + (1|site),
                      data = algae_scaled_fringe)  
  performance::check_model(m.synch.stab.cov.FRsc)
  
# effect of richness on cover
m.sprich.cov.FRsc <- glmer(cover_mean_scale ~ richness_mean_scale + (1|site),
                      data = algae_scaled_fringe)
  performance::check_model(m.sprich.cov.FRsc)
```

Try it with piecewiseSEM

```{r}
# Build SEM
sem.FRsc <- psem(
  m.sprich.synch.FRsc,
  m.synch.stab.cov.FRsc,
  m.sprich.cov.FRsc,
  data = as.data.frame(algae_scaled_fringe)
)

summary(sem.FRsc)
```

Plot

```{r}
plot(sem.FRsc)
```



