---
title: "Supplemental tables to figures"
output: html_document
date: "2025-03-10"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(here) # relative file paths
require(tidyverse) # data wrangling
require(patchwork) # putting plots together
```

# Load data

```{r}
supp_tables <- read_csv(here("data/supplemental_tables.csv"))
```

# Set plotting colors and plotting theme

```{r}
habitat_colours <- c("Fringing" = "#D81B60",
                     "Backreef" = "#FFC107",
                     "Forereef 10m" = "#1E88E5",
                     "Forereef 17m" = "#004D40")

model_themes <-   theme_bw() + 
  theme(axis.title.x = element_text(size=24), 
        axis.title.y = element_text(size=24), 
        axis.text.x = element_text(size=22), 
        axis.text.y = element_text(size=22),
        legend.text = element_text(size=24),
        # add the same size for facet_wrap titles
        strip.text = element_text(size = 24),
        plot.title = element_text(size=26), 
        panel.background = element_blank(), 
        panel.grid.major = element_blank(),  #remove major-grid labels
        panel.grid.minor = element_blank(),  #remove minor-grid labels
        plot.background = element_rect(fill = "white"),
        legend.title = element_blank(),
        legend.position = 'bottom', # move legend to bottom
        legend.box = 'horizontal', # make legend horizontal
        legend.box.background = element_rect(colour = "black"))  # add box around legend
```


# Plot S2

First make the plot-level panel:

```{r}
supp_tables %>% 
  filter(Table == "S2") %>% 
  filter(`Spatial scale` == "Plot-level") %>% 
  # order the habitats by distance from shore
  mutate(Habitat = factor(Habitat, levels = c("Forereef 17m", "Forereef 10m",  "Backreef", "Fringing"))) %>% 
  ggplot(aes(x = Mean, y = Habitat, color = Habitat)) +
  geom_point(size = 6) +
  # add confidence intervals:
  geom_linerange(aes(xmin = Lower_CI, xmax = Upper_CI), linewidth = 2.5) +
  facet_wrap(~ Predictor) +
  scale_colour_manual(values = habitat_colours) +
  xlab("Coefficient") +
  ylab("") +
  ggtitle("(a) Plot-level") +
  model_themes + 
  # add letters denoting significance
  geom_text(aes(x = Upper_CI + 0.03 , y = Habitat, 
                label = Letter), colour = "black", size = 10) +
  theme(legend.position = "none") -> s2.plt.level
```

Then the site-level:

```{r}
supp_tables %>% 
  filter(Table == "S2") %>% 
  filter(`Spatial scale` == "Site-level") %>% 
  # order the habitats by distance from shore
  mutate(Habitat = factor(Habitat, levels = c("Forereef 17m", "Forereef 10m",  "Backreef", "Fringing"))) %>% 
  ggplot(aes(x = Mean, y = Habitat, color = Habitat)) +
  geom_point(size = 6) +
  # add confidence intervals:
  geom_linerange(aes(xmin = Lower_CI, xmax = Upper_CI), linewidth = 2.5) +
  facet_wrap(~ Predictor) +
  scale_colour_manual(values = habitat_colours) +
  xlab("Coefficient") +
  ylab("") +
  ggtitle("(b) Site-level") +
  model_themes + 
  # add letters denoting significance
  geom_text(aes(x = Upper_CI + 0.03 , y = Habitat, 
                label = Letter), colour = "black", size = 10) -> s2.site.level
```

Plot together using `patchwork`

```{r}
pdf(here("output/tableS2toFig.pdf"), height = 12, width = 13)

s2.plt.level / s2.site.level

dev.off()
```

# Plot S3

Taxonomic richness:

```{r}
supp_tables %>% 
  filter(Table == "S3") %>% 
  filter(Predictor == "Taxonomic richness") %>% 
  # order the habitats by distance from shore
  mutate(Habitat = factor(Habitat, levels = c("Forereef 17m", "Forereef 10m",  "Backreef", "Fringing"))) %>% 
  ggplot(aes(x = Mean, y = Habitat, color = Habitat)) +
  geom_point(size = 6) +
  # add confidence intervals:
  geom_linerange(aes(xmin = Lower_CI, xmax = Upper_CI), linewidth = 2.5) +
  facet_wrap(~ Predictor) +
  scale_colour_manual(values = habitat_colours) +
  xlab("Coefficient") +
  ylab("") +
  ggtitle("(a) Taxonomic richness") +
  model_themes + 
  # add letters denoting significance
  geom_text(aes(x = Upper_CI + 0.01 , y = Habitat, 
                label = Letter), colour = "black", size = 10) +
  theme(legend.position = "none") -> s3.taxon
```
 
Functional richness:

```{r}
supp_tables %>% 
  filter(Table == "S3") %>% 
  filter(Predictor == "Functional richness") %>% 
  # order the habitats by distance from shore
  mutate(Habitat = factor(Habitat, levels = c("Forereef 17m", "Forereef 10m",  "Backreef", "Fringing"))) %>% 
  ggplot(aes(x = Mean, y = Habitat, color = Habitat)) +
  geom_point(size = 6) +
  # add confidence intervals:
  geom_linerange(aes(xmin = Lower_CI, xmax = Upper_CI), linewidth = 2.5) +
  scale_colour_manual(values = habitat_colours) +
  xlab("Coefficient") +
  ylab("") +
  ggtitle("(b) Functional richness") +
  model_themes + 
  # add letters denoting significance
  geom_text(aes(x = Upper_CI + 0.01 , y = Habitat, 
                label = Letter), colour = "black", size = 10) +
  theme(legend.position = "none") -> s3.functional
```

Synchrony:

```{r}
supp_tables %>% 
  filter(Table == "S3") %>% 
  filter(Predictor == "Synchrony") %>% 
  # order the habitats by distance from shore
  mutate(Habitat = factor(Habitat, levels = c("Forereef 17m", "Forereef 10m",  "Backreef", "Fringing"))) %>% 
  ggplot(aes(x = Mean, y = Habitat, color = Habitat)) +
  geom_point(size = 6) +
  # add confidence intervals:
  geom_linerange(aes(xmin = Lower_CI, xmax = Upper_CI), linewidth = 2.5) +
  scale_colour_manual(values = habitat_colours) +
  xlab("Coefficient") +
  ylab("") +
  ggtitle("(c) Synchrony") +
  model_themes + 
  # add letters denoting significance
  geom_text(aes(x = Upper_CI + 0.10 , y = Habitat, 
                label = Letter), colour = "black", size = 10) -> s3.synch
```

Combine using `patchwork`:

```{r}
pdf(here("output/tableS3toFig.pdf"), height = 17, width = 11)

s3.taxon / s3.functional / s3.synch

dev.off()
```

# Plot S4

Site-level taxonomic richness

```{r}
supp_tables %>% 
  filter(Table == "S4") %>% 
  filter(Predictor == "Site-level taxonomic-richness") %>% 
  # order the habitats by distance from shore
  mutate(Habitat = factor(Habitat, levels = c("Forereef 17m", "Forereef 10m",  "Backreef", "Fringing"))) %>% 
  ggplot(aes(x = Mean, y = Habitat, color = Habitat)) +
  geom_point(size = 6) +
  # add confidence intervals:
  geom_linerange(aes(xmin = Lower_CI, xmax = Upper_CI), linewidth = 2.5) +
  scale_colour_manual(values = habitat_colours) +
  # if you want the 0 reference:
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 1, color = "gray") +
  xlab("Coefficient") +
  ylab("") +
  ggtitle("(a) Site-level taxonomic-richness") +
  model_themes + 
  # add letters denoting significance
  geom_text(aes(x = Upper_CI + 0.05 , y = Habitat, 
                label = Letter), colour = "black", size = 10) +
  theme(legend.position = "none") -> s4.taxon
```


```{r}
supp_tables %>% 
  filter(Table == "S4") %>% 
  filter(Predictor == "Site-level functional-richness") %>% 
  # order the habitats by distance from shore
  mutate(Habitat = factor(Habitat, levels = c("Forereef 17m", "Forereef 10m",  "Backreef", "Fringing"))) %>% 
  ggplot(aes(x = Mean, y = Habitat, color = Habitat)) +
  geom_point(size = 6) +
  # add confidence intervals:
  geom_linerange(aes(xmin = Lower_CI, xmax = Upper_CI), linewidth = 2.5) +
  scale_colour_manual(values = habitat_colours) +
    # if you want the 0 reference:
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 1, color = "gray") +
  xlab("Coefficient") +
  ylab("") +
  ggtitle("(b) Site-level functional-richness") +
  model_themes + 
  # add letters denoting significance
  geom_text(aes(x = Upper_CI + 0.1 , y = Habitat, 
                label = Letter), colour = "black", size = 10) +
  theme(legend.position = "none") -> s4.functional
```

Site-level species synchrony

```{r}
supp_tables %>% 
  filter(Table == "S4") %>% 
  filter(Predictor == "Site-level species synchrony") %>% 
  # order the habitats by distance from shore
  mutate(Habitat = factor(Habitat, levels = c("Forereef 17m", "Forereef 10m",  "Backreef", "Fringing"))) %>% 
  ggplot(aes(x = Mean, y = Habitat, color = Habitat)) +
  geom_point(size = 6) +
  # add confidence intervals:
  geom_linerange(aes(xmin = Lower_CI, xmax = Upper_CI), linewidth = 2.5) +
  scale_colour_manual(values = habitat_colours) +
    # if you want the 0 reference:
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 1, color = "gray") +
  xlab("Coefficient") +
  ylab("") +
  ggtitle("(c) Site-level species synchrony") +
  model_themes + 
  # add letters denoting significance
  geom_text(aes(x = Upper_CI + 0.4 , y = Habitat, 
                label = Letter), colour = "black", size = 10) +
  theme(legend.position = "none") -> s4.synch
```

Mean plot-level species synchrony

```{r}
supp_tables %>% 
  filter(Table == "S4") %>% 
  filter(Predictor == "Mean plot-level species synchrony") %>% 
  # order the habitats by distance from shore
  mutate(Habitat = factor(Habitat, levels = c("Forereef 17m", "Forereef 10m",  "Backreef", "Fringing"))) %>% 
  ggplot(aes(x = Mean, y = Habitat, color = Habitat)) +
  geom_point(size = 6) +
  # add confidence intervals:
  geom_linerange(aes(xmin = Lower_CI, xmax = Upper_CI), linewidth = 2.5) +
  scale_colour_manual(values = habitat_colours) +
    # if you want the 0 reference:
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 1, color = "gray") +
  xlab("Coefficient") +
  ylab("") +
  ggtitle("(d) Mean plot-level species synchrony") +
  model_themes + 
  # add letters denoting significance
  geom_text(aes(x = Upper_CI + 0.6 , y = Habitat, 
                label = Letter), colour = "black", size = 10) -> s4.plt.synch
```

Spatial synchrony

```{r}
supp_tables %>% 
  filter(Table == "S4") %>% 
  filter(Predictor == "Spatial synchrony") %>% 
  # order the habitats by distance from shore
  mutate(Habitat = factor(Habitat, levels = c("Forereef 17m", "Forereef 10m",  "Backreef", "Fringing"))) %>% 
  ggplot(aes(x = Mean, y = Habitat, color = Habitat)) +
  geom_point(size = 6) +
  # add confidence intervals:
  geom_linerange(aes(xmin = Lower_CI, xmax = Upper_CI), linewidth = 2.5) +
  scale_colour_manual(values = habitat_colours) +
    # if you want the 0 reference:
  geom_vline(xintercept = 0, linetype = "dashed", linewidth = 1, color = "gray") +
  xlab("Coefficient") +
  ylab("") +
  ggtitle("(f) Spatial synchrony") +
  model_themes + 
  # add letters denoting significance
  geom_text(aes(x = Upper_CI + 0.4 , y = Habitat, 
                label = Letter), colour = "black", size = 10) +
  theme(legend.position = "none") -> s4.plt.synch
```

Plot together using `patchwork`

```{r}
pdf(here("output/tableS4toFig.pdf"), height = 15, width = 32)

(s4.taxon + s4.functional + s4.synch) /
  (s4.plt.stab + s4.plt.synch + s4.spat.synch)

dev.off()
```

